Transition path sampling based calculation of free energies of enzymatic reactions

### 2022-04-13 tex/manuscript.tex
added some more text

TPS in particular is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS can be considered as a generalization of the Monte Carlo sampling of the 
configuration space to the space of trajectories where the collected ensemble of 
reactive pathways are real dynamical trajectories compared to articial trajectories
generated using bias forces. The TPS ensemble can be used to obtain kinetic information
such as the rate of the reaction as well as the transition state, reaction
coordinate and the reaction mechanism. TPS has been successfully 
applied to study problems such as enzyme catalysis, protein folding, etc.

Since TPS focuses on collecting only reactive trajectories the information contained 
in the non-reactive trajectories which is essential for describing the 
equilibrium distribution is missing. 
Methods that can be used to simultaneously calculate
the dynamical as well as equilibrium properties of a system within 
a certain set of asssumptions have garnered interest in recent times. \cite{Suarez14JChemTheoryComput10p2658}
Transition interface sampling (TIS) is a variant of TPS which can be used for 
this purpose. \cite{vanErp03JChemPhys118p7762}

A common theme among such methods is to divide the reaction coordinate into
windows within which independent sampling can be performed and interpolated
in a way that can used to calculate equilibrium properties

To calculate equilibrium properties such as the free energy 
of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}

@@

\section{Results and discussion}

@@

\subsection{QMMM simulations}

@@

QMMM simulations were used for the TPS simulations in this work and the 
CHARMM force field \cite{Brooks09JComputChem30p1545} is used for the 
MM portion and the PM3 semiempirical method \cite{Repasky02JComputChem23p1601} 
is used for the QM region.

%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
reaction between adenosine triphosphaste and L-methionine which leads to the 
formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
The mechanism of this reaction is determined to follow a SN$_2$ transition state
with the sulphur atom of the L-methionine as the nucleophile with the 
triphosphate leaving group.

@@

\begin{figure}
\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
We used transition path sampling to generate and
analyze the ensemble of reactive trajectories. As order
parameters that define the reactant and product basins, we
used the donor-hydride and acceptor-hydride distances. The
reactant basin was defined as a donor-hydride distance smaller
than 1.6  and the product basin as an acceptor-hydride
distance smaller than 1.6 . To seed the generation of the path
ensemble, we generated a biased trajectory by applying a
harmonic force between the hydride and the acceptor and
allowed the system to propagate for 250 fs. New trajectories
were generated by the shooting algorithm: we perturbed the
momenta of a randomly chosen slice of a reactive trajectory,
and using the new momenta, we allowed the system to
propagate forward and backward in time for 250 fs. A total of
120 reactive trajectories were generated.
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}

>>>>

TPS in particular is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
complex for large biological systems.
TPS is based on the Monte Carlo importance sampling in the 
space of trajectories where the collected ensemble of 
reactive pathways are real dynamical trajectories generated without the 
application of any bias forces. Thus the TPS ensemble can be used to obtain 
kinetic information such as the rate of the reaction as well as to elucidate 
the transition state, reaction coordinate and the reaction mechanism. 
TPS has been successfully applied to study problems such as enzyme catalysis, 
protein folding, etc.

Calculating equilibrium properties such as the free energy of an enzymatic 
catalysis reaction within TPS needs more attention. 
Since the TPS path ensemble only consists of reactive trajectories, the information from 
the discarded non-reactive trajectories which is required for obtaining the 
equilibrium probability distribution of the order parameter is missing. \cite{Dellago09AdvCompSimAppp167}

Methods that can be used to simultaneously calculate
the dynamical as well as equilibrium properties of a system within 
a certain set of asssumptions have garnered interest in recent times. \cite{Suarez14JChemTheoryComput10p2658}
Transition interface sampling (TIS) is a variant of TPS which can be used for 
this purpose. \cite{vanErp03JChemPhys118p7762}

@@

\section{Computational details}

@@

\subsection{System preparation}

@@

We used the crystal structure of human MAT2A 
bound to methylthioadenosine and di-imido triphosphate (PNPNP)
(7L1A from the protein data bank (PDB)) reported by Niland 
et al. \cite{Niland21Biochem60p791} and the crystal structure of \textit{pv}ADA 
in complex with MT-coformycin (3EWC from the PDB) reported by Ho et 
al. \cite{Ho09Biochemistry48p9618} as the starting points of our two 
simulations. Molecular dynamics simulations within classical and hybrid 
quantum mechanics/molecular mechanics (QMMM) approximations were performed 
using the CHARMM program package and using the CHARMM 36 force 
field \cite{Brooks09JComputChem30p1545} for the MM part and the 
approximate PM3 semiempirical method \cite{Repasky02JComputChem23p1601} for the QM part.

Each protein was solvated within
a 15 $\AA$ sphere using TIP3 waters and neutralized using
potassium ions. We minimized the system with 50 steps of the
Steepest Descent method, followed by 2000 steps of the
adopted Basis Newton-Raphson method. The system was
heated slowly to 300 K for 30 ps, starting with harmonic
constraints except on H atoms and TIP3 waters that were
gradually reduced to force constant 1. We started the
equilibration with 15 ps with harmonic constraints of force
constant 1, which were then removed and the equilibration ran
for 60 ps.

We used transition path sampling to generate and
analyze the ensemble of reactive trajectories. As order
parameters that define the reactant and product basins, we
used the donor-hydride and acceptor-hydride distances. The
reactant basin was defined as a donor-hydride distance smaller
than 1.6  and the product basin as an acceptor-hydride
distance smaller than 1.6 . To seed the generation of the path
ensemble, we generated a biased trajectory by applying a
harmonic force between the hydride and the acceptor and
allowed the system to propagate for 250 fs. New trajectories
were generated by the shooting algorithm: we perturbed the
momenta of a randomly chosen slice of a reactive trajectory,
and using the new momenta, we allowed the system to
propagate forward and backward in time for 250 fs. A total of
120 reactive trajectories were generated.
%------------------------------------------------------------------------------ 
\subsection{Plasmodium vivax adenosine deaminase}
%------------------------------------------------------------------------------

%------------------------------------------------------------------------------ 
\section{Results and discussion}
%------------------------------------------------------------------------------ 
\subsection{Human Methionine Adenosyltransferase 2A}
%------------------------------------------------------------------------------ 
Human methionine adenosyltransferase (MAT2A) is an enzyme that catalyzes the 
reaction between adenosine triphosphaste and L-methionine which leads to the 
formation of S-Adenosylmethionine. \cite{Firestone17JAmChemSoc139p13754,Niland21Biochem60p791} 
The mechanism of this reaction is experimentally determined to follow a 
SN$_2$ transition state with the sulphur atom of the L-methionine as the 
nucleophile with the triphosphate leaving group.

@@

\begin{figure}
\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}

%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}
